---
title: "Day 3 R programming"
output: github_document
date: "2026-05-03"
---


Today, I am learning about the different Data Structures in R. the first of many: Vectors and Matrices.

## 1. Vectors

It is a 1-dimensional list of items. To create a vector, we will use the combine c() function.

```{r to create vectors}
grades = c(90,85,92,83)

names = c("A", "B", "C", "D")

grades
```

## 2. Math operations in Vector

```{r}
grades = grades + 5
grades

sum(grades)
mean(grades)
max(grades)
```

## 3. Extracting data from vectors:

```{r}
names[2]
names[c(1,3)]
grades[c(1,3)]

names[-1]

grades[grades > 90]

```

## 4. Matrices:

It is same like a vector, but it has rows and columns, so it's 2-dimensional. Both vectors and matrices are homogeneous in R. Meaning: each variable will have a list of items of only one data type, not a mix.

we will use the matrix() function for this. It has byrow=TRUE/FALSE -> means to fill the matrix horizontally or vertically.

```{r}
m = matrix(1:12, nrow=4, ncol=3, byrow=TRUE)
m
```