---
title: "Day 17 of R Programming"
output: github_document
date: "2026-05-17"
---

Today, I am learning about how to visualize statistical distributions using `ggplot2`.


```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
```

## 1. Our Practice Dataset
Let's build a mock dataset of 500 Kata tournament scores from athletes practicing two different styles. 

```{r create_data}
# We use set.seed() so the random numbers generate the same way every time we Knit!
set.seed(42) 

kata_data <- tibble(
  athlete_id = 1:500,
  style = sample(c("Shotokan", "Shito-ryu"), 500, replace = TRUE),
  # Generating random scores around an 8.0 average
  score = round(rnorm(500, mean = 8.0, sd = 0.5), 1) 
)

head(kata_data)
```

## 2. Histograms (`geom_histogram`)
A histogram takes a single continuous variable (like `score`) and drops it into "bins" to show how frequently certain numbers appear. It only requires an `x` aesthetic!

```{r histogram}
ggplot(kata_data, aes(x = score)) +
  # We define 'binwidth' to determine how wide each bucket of scores is
  geom_histogram(binwidth = 0.1, fill = "steelblue", color = "black") +
  labs(
    title = "Distribution of Kata Scores",
    x = "Final Score",
    y = "Number of Athletes"
  ) +
  theme_minimal()
```

## 3. Density Plots (`geom_density`)

```{r density_plot}
ggplot(kata_data, aes(x = score, fill = style)) +
  geom_density(alpha = 0.5) + 
  labs(
    title = "Score Density by Karate Style",
    x = "Score",
    y = "Density"
  ) +
  theme_minimal()
```

## 4. Boxplots (`geom_boxplot`)

```{r boxplot}
ggplot(kata_data, aes(x = style, y = score, fill = style)) +
  geom_boxplot() +
  labs(
    title = "Score Variation and Outliers",
    subtitle = "Comparing scoring consistency between styles",
    x = "Kata Style",
    y = "Final Score"
  ) +
  theme_bw() +
  theme(legend.position = "none") 
```

## 5. Layering Geometries: Violin Plots
A `geom_violin` is basically a mirrored density plot.

```{r violin_plot}
ggplot(kata_data, aes(x = style, y = score, fill = style)) +
  geom_violin(alpha = 0.4) +
  geom_boxplot(width = 0.1, fill = "white", color = "black") + 
  labs(title = "Violin Plot with Embedded Boxplot") +
  theme_minimal() +
  theme(legend.position = "none")
```